Baseline centering problem

### 2018-12-19 paper/tangotd.tex

The Newton update/middle Bellman equation is
\begin{equation}
Z\gets (1+\eta)Z-\eta Z^2+\eta \gamma ZPZ
\end{equation}
which yields (after applying to $\1_{s_1}$ on the left and $\1_{s_2}/\rho(s_2)$
on the right)
\begin{multline}
Z(s_1,s_2)\gets (1+\eta)Z(s_1,s_2)-\eta \E_{s_3\sim \rho}\,
Z(s_1,s_3)Z(s_3,s_2)\,+\\\eta \gamma \E_{s_3\sim\rho,\, s'_3\sim
P(s'_3|s_3)}\,
Z(s_1,s_3)Z(s'_3,s_2)
\end{multline}
which is perfectly samplable. We can simulate the assignment
$Z(s_1,s_2)\gets...$ via gradient descent on the parameters of $Z$.

@@

Ideally, given the transition $s_3\to s'_3$, I'd like to sample $s_1$ and
$s_2$ proportionally to $Z$ (would cancel the $Z$ factors, and avoid
applying the equation to states that have nothing to do with
$s_3$—although the $Z$ factors are different for the three terms...). But
I don't know how to build an efficient sampler from $Z(s_3,\cdot)$.
Something like Metropolis? but this function $Z$ is not constant...
On-policy, sampling from $Z(\cdot,s_3)$ is easy: just sample one of the
recent states with geometric rates $\gamma$. This should divide the
equation by $Z(s_1,s_3)$ (up to some $\gamma$-dependent factor). I'm not
sure what to do with the resulting $Z(s_1,s_2)/Z(s_1,s_3)$. On-policy, another
option is to sample triplets $(s_1,s_3,s_2)$ that are some time apart on
the trajectory, e.g., starting with $s_3$ uniform along the past
trajectory, and moving back for $s_1$ and forward for $s_2$ (with factors
$\gamma$). Looks a lot like eligibility traces?

>>>>

The Newton update/middle Bellman equation is
\begin{equation}
Z\gets (1+\eta)Z-\eta Z^2+\eta \gamma ZPZ
\end{equation}
which yields (after applying to $\1_{s_1}$ on the left and $\1_{s_2}/\rho(s_2)$
on the right)
\begin{multline}
Z(s_1,s_2)\gets (1+\eta)Z(s_1,s_2)-\eta \E_{s_3\sim \rho}\,
Z(s_1,s_3)Z(s_3,s_2)\,+\\\eta \gamma \E_{s_3\sim\rho,\, s'_3\sim
P(s'_3|s_3)}\,
Z(s_1,s_3)Z(s'_3,s_2)
\end{multline}
which is perfectly samplable. We can simulate the assignment
$Z(s_1,s_2)\gets...$ via gradient descent on the parameters of
$Z(s_1,s_2)$. (Or via double sampling, which is not problematic here!?)

@@

Ideally, given the transition $s_3\to s'_3$, I'd like to sample $s_1$ and
$s_2$ proportionally to $Z$ (would cancel the $Z$ factors, and avoid
applying the equation to states that have nothing to do with
$s_3$—although the $Z$ factors are different for the three terms...). But
I don't know how to build an efficient sampler from $Z(s_3,\cdot)$.
Something like Metropolis? but this function $Z$ is not constant...

On-policy, sampling from $Z(\cdot,s_3)$ is easy: just sample one of the
recent states with geometric rates $\gamma$. This should divide the
right part of the equation by $Z(s_1,s_3)$ (up to some $\gamma$-dependent
factor). Thus we would sample two states $s_1,s_2$ along the trajectory,
sample a state $s_3$ in between on the trajectory from $s_1$, and (with
the right $\gamma$ factors) we would get a term with $Z(s_1,s_2)$, a term
with $Z(s_3,s_2)$ and a term with $Z(s'_3,s_2)$. This looks suspiciously
like a large-scale Bellman from $s_1$... Variance could be higher since
we don't use the model $Z$ and use sampling instead. To be tested.
On-policy, another
option is to sample triplets $(s_1,s_3,s_2)$ that are some time apart on
the trajectory, e.g., starting with $s_3$ uniform along the past
trajectory, and moving back for $s_1$ and forward for $s_2$ (with factors
$\gamma$). Looks a lot like eligibility traces?
